\hypertarget{sec:methodology}{%
\chapter{Verification Methodology}\label{sec:methodology}}

Formal verification takes many forms, with many techniques and tools at our disposal, and choosing the correct verification tool is of utmost importance.
In the past, \Abrev{BAN logic} \autocite{burrows1989logic} was the tool of choice for verifying cryptographic protocols.
Unfortunately, protocol components verified by \Abrev{BAN logic} were later found to have undiscovered vunlerabilities \autocite{snekkenes1991exploring}.
BAN was subsequently critiqued \autocite{boyd1993limitation} for not being able to fully verify crucial aspects of cryptographic protocols.
Because of such concerns, results and conclusions from verification techniques using \Abrev{BAN logic} stating that no attacks exist can be a false-positive.
The kind false-positive verification of security that \Abrev{BAN logic} provided could cause the very lapses in post-adoption security that the verification had attempted to preempt.
\Abrev{BAN logic} has since fallen out of favor and instead been replaced by model checking techniques \autocite{holzmann1991design} and theorem proving \autocite{andrews1980axiomatic}, which have been able to produce attacks on cryptographic protocols which BAN techniques incorrectly verified as secure.

Theorem proving, when possible, is the preferred method of cryptographic protocol verification \autocite{kobeissi2020verifpal}.
Simple communication protocols have had success with verificiation via formal description and theroem prover \autocite{berry1987towards, cheng1990formal}.
Generally, modern protocols are constructed in too complex a manner for a theorem prover to verify, with automated theorem proving still being an area of active research \autocite{xiong2020smartverif, bhargavan2021text}.
When defining a protocol in terms compatible with modern theorem provers issues of termination checking, proof tactic selection, and SMT limitations greatly impede computer automated verification \autocite{hassan2021automated}.
Human guidance is still required to direct theorem provers towards the verification goal.
Some domain specific theroem provers, such as Tamarin \autocite{basin2017symbolically}, exist for cyrptography and communication protocols to reduce the required human effort.
The specialized and time-consuming nature of human effort required is a major consideration when deciding between utilizing a theorem prover and a model checker.
Contrasted with theorem provers, model checkers are highly valued for the automation and reduced human effort they provide, with the work presented here being no exception.

This work presents the verification of the TreeKEM protocol's \Abrev{FSU} and \Abrev{PCS} security guarantees.
To verify these security guarantees, the \CGKAsec\ is used as an abstraction of the TreeKEM protocol, as well as all other \CGKAdef-conforming protocols.
The \CGKAsec\ will be modeled as an explicit \Abrev{FSM}, representing all possible TreeKEM execution states.
More precisely \CGKAsec\ will be modeled as a \Abrev{LTS} \autocite{keller1976formal}, which includes informational labels for each state in the \Abrev{FSM}.
However, much of the literature related to the chosen model checker uses the nomenclature of \Abrev{FSM} instead of \Abrev{LTS}, and this work will follow suit.
The security guarantees will be encoded as \Abrev{LTL} statements \autocite{4567924}.
Then explicit state model checking techniques will be applied to verify that the \Abrev{LTL} encodings of \Abrev{FSU} and \Abrev{PCS} are maintained in every possible state of the \CGKAsec.

There is one last important detail the verification.
Recall that what differentiates \Abrev{SM} from \Abrev{SGM} is that \Abrev{SGM} involves two or more parties.
There is no theoretical limit to the number of parties the TreeKEM protocol supports.
However, we also stated that model checking represents all possible states of a protocol as a finite graph of transitions.
Here we have a potential problem, where TreeKEM allows an unbounded number of participants, producing an infinite number of possible states, while model checkers require a finite number of states.
The IETF states that MLS in practice should support approximately $50,000$ participants \autocite{Omara2020}, which is a finite, yet this very large number is problematic for the chosen methodology of explicit state model checking.
This work parameterizes the model in terms of \(T\) and \(N\) from the definition of \emph{Non-adaptive} \((T, C, N, \textbf{\texttt{P}}, \epsilon)\) \emph{CGKA Security}and limits the scope of verification to within the selected parameters.
In practice, this means the results presented for a specified \Tmax\ and \Nmax\ only provide verification of security against \((T, C, N)\)-attackers for all \(T \in [1, \Tmax]\), \(C \in [1, T]\) and \(N \in [2, \Nmax]\).

Referring back to the verification hypotheses in Table\ \ref{tab:verification-hypotheses}, any outcome which falsifies a hypothesis is an interesting result.
If a counter-example is produced violating \Abrev{PCS} for either version of the protocol, then a previously unknown security defect for TreeKEM has been discovered.
If a counter-example is produced violating \Abrev{FSU} for the revised \VersionTwo, this also discovers a security defect.
The original \VersionOne\ has been shown to have a \Abrev{FSU} deficiency for a \((12, 12, 8)\)-attacker \(\mathcal{A}\) \autocite{alwen2020security}.
Hence a successful \Abrev{FSU} verification for the original \VersionOne\ with parameters corresponding to \(\mathcal{A}\) indicates either an in the model.
It has not been stated whether the example provided by \autocite{alwen2020security} does not state that this example is minimal.
Locating the minimal \(T\), \(C\), and \(N\) values for which the vulnerability manifests in the \VersionOne version.
In order to ensure at least one observation of the \Abrev{FSU} deficiency of the original \VersionOne, the bounds are set to \(12 = \Tmax\) and \(8 = \Nmax\).
The observed verification results are tabulated in Appendix\ \ref{apx:Observations} and discussed in Chapter\ \ref{results}.


\hypertarget{sec:explicit-state-model-checking}{%
\section{Explicit state model checking}\label{sec:explicit-state-model-checking}}

Model checking exists as the preferred alternative to \Abrev{BAN logic} techniques for verifying cryptographic protocols \autocite{kacprzak2006comparing, lomuscio2007verification, van2004symbolic}.
Model checking represents all possible states of protocol as an \Abrev{FSM} \autocite{clarke1981design}.
There exist one or more start states in the transition graph representing possible starting condiitions.
Transitions in the graph uniquely encode the possible changes in model state.
The model checker then asserts that one or more properties holds for the protocol by traversing through all combinations of transitions through the graph \autocite{Vardi1986}.
If a property holds, the model checker will assert that there is no possible set of transitions through states that exist in the protocol which violate the specified property.
If the property does not hold, the model checker will produce a counter-example sequence of transitions through possible states showing how the specified property can be violated.
This method of verifying cryptographic protocols is very attractive as the verification result is either an exhaustive proof of security guarantees or a cryptographic attack which can be analyzed and corrected.
This constructive proof methodology offered by explicit state model checking, the guarantee of a falsifying counterexample if and only if one exists, is of paramount important for correcting defects within the modeled protocol.

The principle drawback of explicit state model checking is the immensity of the state-space as the complexity of the model grows.
Each possible state in the model is uniquely represented and all transitions for all states are computed during verification.
Controlling memory usage to represent and traverse the search-space without greatly sacrificing the speed of verification remains a problem with limited tractability.
For TreeKEM, the parameters \(T\) and \(N\) are supplied to the model.
The model is then verified by exhaustively searching all possible protocol states over \(T\) epochs and with the total unique group members ranging from \(2\) to \(N\).
Incrementing the \(T\) parameter while holding \(N\) constant, generally, increases the state-space by \(2\) orders of magnitude.
However, incrementing the \(N\) parameter while holding \(T\) constant, only increases the state-space by \(\frac{1}{2}\) an orders of magnitude.
Due to the constraints engineering constraints presented by explicit state model checking, the verification work presented will only consider the combinatorial collection of all models parameterized by \(T\) and \(N\), where \(T \in [1, \Tmax]\) and \(N \in [2, \Nmax]\).


\hypertarget{spin}{%
\section{Spin}\label{spin}}

Among the many options of model checkers, I have chosen use the SPIN model checker \autocite{godefroid1990using, holzmann1997model}.
SPIN provides several important features for our verification goals.
The main difficulty with model checking is the size of the number of states and transitions \autocite{burch1992symbolic}.
Since model checking works by checking every possible combination of states and transitions, the size of the state-space directly impacts the feasibility.
SPIN provides some solutions to making this problem tractable, the culmination of over three decades of research and development.
First, SPIN supports ``on-the-fly'' verification \autocite{rudin1987limits}, which means that SPIN does not need to consider the entire set of states and transitions at the same time, but instead can work on just subsets until the whole set is verified.
Second, SPIN also uses a ``partial ordering'' of the states and transitions \autocite{shannon1958note} to reduce the number of subsets it needs to verify before being certain that a property holds for the whole set of states and transitions.
Both ``on-the-fly'' verification and ``partial order'' reduction are enabled by default as they can be used in conjunction with minimal drawbacks \autocite{valmari1993fly, peled1994combining, holzmann1995improvement, couvreur1999fly}.
Consideration was made to use TopSpin \autocite{DonaldsonM_AMAST2006}, a PROMELA pre-processing plugin for SPIN which performs symmetry reduction as well.
Unfortunately, this theoretically viable technique is not practically usable as the plugin has not been maintained in over a decade.
Additionally, SPIN supports numerous compile time directive which alter the state-space representation, search trajectories, and time-memory trade-offs.
Because of these (and many more) techniques, SPIN allows efficiently working with a large state-space that TreeKEM produces when modeled with PROMELA.\@


\hypertarget{sec:promela}{%
\section{Promela}\label{sec:promela}}

With SPIN selected as the model checker, the obvious modeling language in which to encode the TreeKEM protocol is the Protocol Meta Language (PROMELA) \autocite{holzmann1980basic, holzmann1990design}.
The key language feature of PROMELA which makes it so conducive to explicit state model checking is that *all* branching -- conditional executaion, looping, and randomness -- are evaluated nondeterministically.
Heance, PROMELA is used to produce all possible states and transitions of TreeKEM, we must supply that representation to a model checker.
It exists as a precise and convenient specification for defining all possible states and transitions for TreeKEM.\@
By specifying the protocol in PROMELA, one can allow the PROMELA ``translator: to take a high-level, human readable description of TreeKEM and produce a much larger machine readable list with all the possible states and transitions in TreeKEM.\@
Manually defining all these states and transitions would take an inordinate amount of time and is prone to errors.
Producing an exhaustive and correct representation of TreeKEM's possible states and transitions is one aspect of the verification process, at which machines are exceptionally reliable, while humans are not.

The language of PROMELA is similar to the venerable C programming language.
The similarity is not accidental, as it allows for the clear translation between a (perhaps the most) popular systems programming language and the modeling language.
PROMELA supports mutable variable assignment, treating all variables as a finite and linearly ordered collection of bits.
Assignments mutate an implicit global state.
Variables can take the form of Boolean values, signed or unsigned numbers of varying bit widths, type-checked enumerations, or multi-dimensional arrays.
PROMELA also provides looping constructs and an inlining of definition, similar to an inlined function call in C, allowing for code reuse.
Examples of usages is dipicted in Figure\ \ref{fig:determinism}.
These familiar C-like features of PROMELA are evaluated deterministically and linearly within the execution of a model defined in PROMELA.\@
However, PROMELA also differs from C in important ways which facilitate explicit state graph construction and exhaustive search.

\begin{figure}
\centering
\caption{\label{fig:determinism} PROMELA example --- deterministic operations}
\begin{verbatim}
// type  identifier   value    range
bool     BitVar     = true; // [0, 1]
byte     MyByte     = 12;   // [-128, 127]
unsigned NonNeg : 6 = 42;   // [0, 63]
int      AnArray[12];       // range of int * 12

// type-checked enumeration with 3 values
mtype:OK = { Pass, Fail, Unknown };

// Looping construct
byte i;
for ( i : 0 .. MyByte ){
    AnArray[i] = 8;
}

// Inline code reuse with parameter aliasing
inline CodeReuse( input ) {
  input = input + MyByte;
  AnArray[0] = input;
}
CodeReuse( MyByte );
CodeReuse( NonNeg );
\end{verbatim}
\end{figure}


PROMELA only provides non-deterministic conditional constructs.
An \texttt{if} statement, shown in Figure\ \ref{fig:choice}, is supplied with one or more logical guards containing PROMELA expression which evaluates to a Boolean value and an associated (possibly empty) sequence PROMELA expressions for each guard.
The model non-deterministically selects a guard which evaluates to \texttt{true} and follows the code path of the associated sequence of expressions.
Once a the non-deterministically selected code path completes execution, execution resumes at the end of the \texttt{if} statement.
Another similar non-deterministic conditional construct is provided by PROMELA.\@
A \texttt{do} statement,statement, shown in Figure\ \ref{fig:repetition}, is also supplied with one or more pairs of logical guards and expression sequences.
However, once a the non-deterministically selected code path of the \texttt{do} statement completes execution, execution resumes at the beginning of the \texttt{do} statement.
Execution only resumes at the end of a \texttt{do} block when one of the non-deterministically selected code paths contains a \texttt{break} statement.
This definition of the \texttt{do} statement provides nondeterministic looping behavior.
The \texttt{if} statement permits identical behavior to C \texttt{if/else} statements if all guards are mutually exclusive.
The \texttt{do} statement permits identical behavior to a while loop with correctly constructed guards.
However, the non-deterministic nature of these conditional constructs provide much more expressive power within the model.

\begin{figure}
\centering
\caption{\label{fig:choice}PROMELA example --- non-deterministic choice via \texttt{if}}
\begin{verbatim}
BitVar = true;
NonNeg = 8;
// Either 2nd or 3rd execution path is possible
if
:: !BitVar && NonNeg < 32 -> NonNeg = NonNeg + 1;
:: BitVar                 -> BitVar = BitVar = false;
:: NonNeg < 32            -> BitVar = ~BitVar;
:: else                   -> NonNeg = 0;
fi
\end{verbatim}
\end{figure}


\begin{figure}
\centering
\caption{\label{fig:repetition}PROMELA example --- non-deterministic repetition via \texttt{do}}
\begin{verbatim}
BitVar = true;
NonNeg = 8;
// Loop ends after at least 24 iterations
do
:: !BitVar && NonNeg < 32 -> NonNeg = NonNeg + 1;
:: BitVar                 -> BitVar = BitVar = false;
:: NonNeg <  32           -> BitVar = ~BitVar;
:: NonNeg >= 32           -> break;
od
\end{verbatim}
\end{figure}

There exists one final nondeterministic construct provided by PROMELA.\@
The \texttt{select} statement, shown in Figure\ \ref{fig:selection}, takes a contiguous, inclusive range of values and non-deterministically selects a value from within the range, assigning the result to a specified variable.
This replaces the notion of a random number generator within the model.
Explicit state model checking requires that all possible states within the model be explicitly enumerated.
Randomness is replaced by a range of nondeterministic values.
During model checking all possible branches of the \texttt{if}, \texttt{do}, and \texttt{select} constructs are explored, creating a wide array of state transitions for exhaustive search.
The expressive power provided by PROMELA's \texttt{if}, \texttt{do}, and \texttt{select} statements allows for conveniently modeling randomness and alternation in verifiably sound manner.

\begin{figure}
\centering
\caption{\label{fig:selection}PROMELA example --- non-determinism via \texttt{select}}
\begin{verbatim}
byte RandomArray[16];
byte i;
for ( i : 0 .. 15 ) {
  byte sample;
  select ( sample : 0 .. 7 );
  RandomArray[i] = sample;
}
// RandomArray now has one of 8 * 16 = 128 possible states.
// All 128 possible RandomArray states will be explored.
\end{verbatim}
\end{figure}

A major difference from other C-like languages is PROMELA's support of concurrency.
A model may include multiple processes, with no restrictions as to when and how the processes are created or terminated.
The model of multiple processes considers the state transitions through each process being interleaved in any permissible permutation during the execution of the model.
To further support modeling concurrency, PROMELA provides an \texttt{atomic} block, shown in Figure\ \ref{fig:atomic}, for indicating that a sequence of expressions cannot be interleaved between processes.

\begin{figure}
\centering
\caption{\label{fig:atomic}PROMELA example --- \texttt{atomic} block}
\begin{verbatim}
byte SharedArray[32];
int  sum = 0; 
byte i;
for ( i : 0 .. 15 ) {
  atomic {
    byte value = RandomArray[i];
    sum = sum + value;
    RandomArray[i] = sum;
  }
}
// Each step of the array scan happens atomically.
\end{verbatim}
\end{figure}


Similarly, PROMELA supports a \texttt{d\_step} block as shown in Figure\ \ref{fig:dstep}, which allows specifying that a sequence of expressions are deterministic within the model and should be considered a single state within the generated state-space transition graph.
Unlike an \texttt{atomic} block, a \texttt{d\_step} block cannot include any nondeterministic constructs.
Judicious use of \texttt{d\_step} can greatly simplify the state-space search as well as improve the comprehensibility of any counterexamples produced by the model checker.

\begin{figure}
\centering
\caption{\label{fig:dstep}PROMELA example --- \texttt{d\_step} block}
\begin{verbatim}
// All steps to assign the value to dim are collapsed.
// There is a single state transition from dim being unassigned
// to the value of dim being computed and finally assigned.
byte dim;
d_step {
  byte m = 10;
  byte n = 12;
  byte t = m * n; 
  t = t * t;
  t = t + t;
  dim = t;
}
\end{verbatim}
\end{figure}

Lastly, PROMELA permits annotating a state with a label.
The labels can the be utilized in a handful of ways to support verification.
Labels prefixed with \texttt{end} identify the set of valid end states in which the program may terminate.
Labels prefixed with \texttt{accept} identify states which the program cannot transition through infinitely often if it executes infinitely.
Labels prefixed with \texttt{progress} identify states which the program must transition through infinitely often if it executes infinitely.
Finally, a labeled state may be referenced in an \Abrev{LTL} predicate.
Examples usages of state labels are illistrated in Figure\ \ref{fig:state-label}.

\begin{figure}
\caption{\label{fig:state-label}PROMELA example --- state labels}
\begin{verbatim}
byte NumArray[6];
int  prod = 0;
byte i = 0;
do
:: i >= 5 -> { 
accept_example: break;
}
:: else -> {
progress_example: i = i + 1
  prod = prod * RandomArray[i%6];
}
end_example: i = prod;
\end{verbatim}
\end{figure}


\hypertarget{sec:linear-temporal-logic}{%
\section{Linear Temporal Logic}\label{sec:linear-temporal-logic}}

Correctly specifying the properties which must hold for the cryptographic protocol is one of the most important aspects of model checking techniques.
Fortunately, \Abrev{LTL} \autocite{4567924} is an almost perfectly tailored language for representing the security guarantees at the CGKA abstraction level.
Additionally, \Abrev{LTL} usage in conjunction with an explicit state model checker is a well supported and researched technique \autocite{clarke1994another, biere1999symbolic, bloem1999efficient}.
\Abrev{LTL} possesses all the logical operators of the two-element Boolean algebra with the addition of four temporal operators.
The temporal operators are presented below as a short had reference for formulas described in Section\ \ref{#sec:LTL-security}.

\[
\begin{aligned}[t]
                 \bigcirc & \varphi & \!\!\!\!\textrm{:} & \;\;\textbf{``Next''}     & \varphi\textrm{ holds at the next state} \hfill~ &\\
                 \Diamond & \varphi & \!\!\!\!\textrm{:} & \;\;\textbf{``Future''}   & \varphi\textrm{ eventually holds, in either the current or future state} \hfill~ &\\
                 \Box     & \varphi & \!\!\!\!\textrm{:} & \;\;\textbf{``Globally''} & \varphi\textrm{ holds at the current state and all future states} \hfill~ &\\
  \psi \;{\mathcal {U}}\, & \varphi & \!\!\!\!\textrm{:} & \;\;\textbf{``Until''}    & \varphi\textrm{ eventually holds and $\psi$ holds at every state until then} \hfill~ &\\
\end{aligned}
\]

An example of \Abrev{LTL}'s expressive power is that it allows for stating that a property must always hold, eventually hold, hold until another property holds, holds in the next state, never holds, or any combination of these.
Often, properties are classified as either \emph{safety}, \emph{liveness}, or \emph{fairness}; though more nuanced which do not adhere the these categories are also expressible with \Abrev{LTL}.
A safety property ensures that an undesirable event never occurs \((\, \Box(\,\neg\varphi \,) \,)\).
A liveness property requires progress towards an event \((\, \Box(\, \varphi \implies \Diamond(\,\psi \,) \,) \,)\).
A fairness property ensures that a desirable even continues to happen \((\, \Box(\, \Diamond(\,\varphi \,) \,) \,)\).
The usage of \Abrev{LTL} allows a precise definition of what must occur in specific segments of time during the the protocol's transition through possible states.
\Abrev{LTL} possesses just enough expressive power to specify the FS and PCS security guarantees, but no more than that.
This perfect balance of expressiveness is important, because the more expressive the logical framework we choose to define our security guarantees, the less efficient the available verification techniques will be.
